\documentclass[12pt]{article}

% House style preamble: central symlink only.
\input{.house-style/preamble.tex}

% --- Anonymization toggle for double-blind review ---
\newif\ifblind
\ifdefined\blindbuild \blindtrue \else \blindfalse \fi

\hypersetup{
  pdftitle={Beyond concordance: External validation of probabilistic grammar in the English dative alternation},
  pdfkeywords={dative alternation, probabilistic grammar, external validation, cross-corpus transport, calibration, projectibility, corpus linguistics}
}
\ifblind\hypersetup{pdfauthor={}}\fi

\title{Beyond concordance\\
{\large External validation of probabilistic grammar in the English dative alternation}}
\ifblind
  \author{}
\else
  \author{Brett Reynolds \orcidlink{0000-0003-0073-7195}%
  \thanks{Contact: \href{mailto:brett.reynolds@humber.ca}{brett.reynolds@humber.ca}}\\
  Humber Polytechnic \& University of Toronto}
\fi
\date{June 2026}

\begin{document}
\maketitle

\begin{abstract}
This paper asks how far a corpus-based model of the English dative alternation
can be carried beyond the corpus that produced it. If a model learned from
the American data analysed by \textcite{bresnan2007} predicts whether speakers
choose \mention{give Kim the book} or \mention{give the book to Kim}, does it
also help with later British conversation, for instance in the Spoken British
National Corpus 2014 (BNC2014)? Earlier studies show that dative models trained on one variety
can rank or classify tokens in another. I ask the stricter question: whether the
probabilities remain reliable, which parts of the conditioning carry over, and
what that production probability can say about grammar. I freeze a
\texttt{languageR} model and score it on BNC2014 with probability scoring,
calibration, a BNC2014-trained comparison on the same rows, predictor-group
ablation, missing-data sensitivity, and an explicit denominator. The source
model ranks BNC2014 choices well, mainly because length, animacy, definiteness,
and pronominality carry over better than verb base rates. But calibration,
coefficient checks, and the BNC2014-trained comparison expose patterns that
ranking alone hides. The model's probability is a conditional production
probability among annotated alternation tokens, not a probability of
grammaticality. It supports prediction for the shared high-frequency dative core
while leaving grammatical but unattested forms to other evidence.
\end{abstract}

\noindent\textbf{Keywords:} dative alternation; probabilistic grammar; external validation; cross-corpus transport; calibration; projectibility; corpus linguistics

\input{sections/01-problem}
\input{sections/02-original-target}
\input{sections/03-data-and-replication}
\input{sections/04-modern-modelling}
\input{sections/05-generalization-and-acceptability}
\input{sections/06-grammatical-possibility}
\input{sections/07-conclusion}

\section*{Data and code availability}

\ifblind
The analysis code and derived files are in a public repository, withheld here
for double-blind review and available to the editors on request; a
de-anonymized link will accompany the camera-ready version.
\else
The public working repository is
\href{https://github.com/BrettRey/bresnan-dative-alternation-reanalysis}{\texttt{BrettRey/bresnan-dative-alternation-reanalysis}}.
The exact submitted version will be archived as a tagged release or
identifier-bearing repository snapshot.
\fi
The analysis scripts fetch public datasets to temporary locations
and validate downloaded files by their MD5 (message-digest~5) checksums.

\begin{center}
\small
\begin{tabular}{ll}
\toprule
File & MD5 \\
\midrule
BNC2014 CSV & \texttt{1c041dbcb635855f3d8f4be9f3e1fced} \\
DAIS item CSV & \texttt{46af310f1633b8784f897e4482faed84} \\
DAIS cleaned-judgement ZIP & \texttt{e7b3ac6f2dd8e85bc93e6e20fc0f6ec1} \\
\bottomrule
\end{tabular}
\end{center}

Derived files and the source-verification note are stored in the repository;
the README gives their paths. The principal reproduction commands are:

\begin{quote}
\small
\texttt{Rscript analysis/08\_dais\_acceptability\_bridge.R}\\
\texttt{Rscript analysis/09\_bnc2014\_metadata\_scope.R}\\
\texttt{Rscript analysis/10\_bnc2014\_paired\_transport\_cv.R}\\
\texttt{python3 analysis/06\_denominator\_and\_figure\_candidates.py}\\
\texttt{make}
\end{quote}

The cross-validation and bootstrap seed is \texttt{20260625}. Session
information is written to a derived text file. Raw DAIS files aren't committed.

\section*{Acknowledgements}

OpenAI Codex and Anthropic Claude assisted with drafting, coding, and review.
They weren't used to create or manipulate submitted figures, images, or
research data. All analysis decisions, source checks, and final text remain the
author's responsibility.

\clearpage
\printbibliography

\clearpage
\input{sections/08-supplementary-dais}

\end{document}
