\section{Data reproduction and reanalysis}

The analysis starts with \texttt{languageR::dative}, distributed with the
Comprehensive R Archive Network (CRAN) \texttt{languageR} package
\citep{BaayenShafaeiBajestan2025languageR}. The source tarball supplies the
packaged \texttt{dative.rda} object; the analysis scripts load it and write
derived summaries.

The package documentation identifies the sources as Switchboard and the
Treebank Wall Street Journal collection. Those sources place the original data
in American English: Switchboard is telephone conversation among speakers from
the United States, developed in 1990--1991
\citep{GodfreyHolliman1993Switchboard1}, and the Treebank Wall Street Journal
component is late-1980s American news and financial writing
\citep{MarcusSantoriniMarcinkiewiczTaylor1999Treebank3,CharniakEtAl2000BLLIPWSJ}.

In the packaged dataset, there are 3,263 rows and 15 columns. The response
records recipient realization, with 2,414 NP realizations and 849 PP
realizations. The dataset contains 75 verb levels and two modalities: 2,360
spoken tokens and 903 written tokens. Speaker identifiers are missing exactly
for the written rows.

For the first model sequence, the response stays binary: NP is coded as 1 and PP as
0. The initial split is verb-stratified, with a fixed seed
(\texttt{20260623}), so every verb observed in the test set is also represented
in training. That split gives 2,637 training rows and 626 test rows.

The denominator is narrow. These rows are already tokens in which a
dative recipient was realized as NP or PP. They aren't all contexts in which a
transfer event could have been expressed, omitted, paraphrased, or packaged in
some other construction. The models estimate conditional choice within an
attested alternation frame: given that a token enters the dative-alternation
sample, which realization appears?

Initial analyses are deliberately plain. A marginal model sets the baseline.
Structured models then add modality, length, verb, and the non-verb predictors.
Negative-control and simulation checks accompany them: a scrambled-label check
and fake-data checks. These tests ask whether the fitted structure beats chance,
not whether it proves a grammatical theory.

A second empirical source is the Spoken British National Corpus 2014 (BNC2014)
dative dataset
\citep{JensetMcGillivray2018BNCData,JensetMcGillivray2019Dative}. The verified
Figshare comma-separated values (CSV) file has 1,839 parsed data rows and 44
columns under a Creative Commons Attribution 4.0 International (CC BY 4.0)
licence.

The BNC2014 dative file comes from the Early-Access Subset of the Spoken
BNC2014, whose spontaneous conversations were recorded in 2012--2014 and
contain speaker and context metadata.

BNC2014 serves as a transport target rather than a second copy of
\texttt{languageR::dative}. Its outcome field is \texttt{Pattern}, where the
verb--noun--noun code \texttt{VNN} maps to NP realization and the
verb--noun--prepositional-phrase code \texttt{VNPP} maps to PP realization.
The six shared verbs are \mention{give}, \mention{send}, \mention{show},
\mention{sell}, \mention{offer}, and \mention{lend}. All six occur in the
packaged \texttt{languageR} dataset.
