\section{Modern modelling update}

Does modern modelling change Bresnan et al.'s linguistic conclusion, or only
repackage it statistically? Here, only the statistical packaging changes. The
table uses three diagnostics to make that answer visible: log loss, area under
the receiver operating characteristic curve (AUC), and a count of fixed-effect
terms.

Two scores carry the paper, so it's worth fixing what they measure. Log loss
rewards probabilities that are both confident and correct and punishes confident
mistakes; it's lower-is-better and reaches zero only for perfect prediction. AUC
is the probability that a randomly chosen NP token is scored above a randomly
chosen PP token, so 0.5 is chance and 1 is perfect ranking. It's the measure
earlier dative-transport studies report as a concordance index or \(C\); the two
names denote one quantity. AUC captures only \term{discrimination}, the ranking
of tokens; log loss also responds to \term{calibration}, whether the predicted
probabilities match observed rates. Those are separate achievements, which is
why both are reported.

All models predict whether the recipient is realized as NP rather than PP. They
are logistic models with R treatment contrasts and no interactions. The
marginal null uses only the training-set NP rate, so it's the base-rate
benchmark. The non-verb model adds modality, semantic class, length, animacy,
definiteness, pronominality, and accessibility, but omits verb identity. It asks
whether the classic non-lexical profile predicts realization without lexical
information.

Two lexical models then ask whether verb-specific biases add anything. The
fixed-verb model gives each verb its own coefficient, so each verb can have a
different baseline preference for NP versus PP. The hierarchical model gives
each verb a varying intercept drawn from a common verb distribution. That's
partial pooling: rare verbs are pulled toward the overall verb mean, while
better-attested verbs can differ more. If this model matches fixed verb effects,
the lexical signal survives regularization.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{lrrr}
\toprule
Model & Test log loss & Test AUC & Fixed terms \\
\midrule
Marginal null & 0.554 & 0.500 & 1 \\
Non-verb main model & 0.271 & 0.932 & 18 \\
Fixed-verb full model & 0.234 & 0.951 & 92 \\
Hierarchical verb-intercept model & 0.233 & 0.952 & 18 \\
\bottomrule
\end{tabular}
\caption{Held-out \texttt{languageR::dative} results for the main model sequence.}
\label{tab:languager-results}
\end{table}

The non-verb model already predicts held-out production choices well. Adding
fixed verb effects improves test log loss from 0.271 to 0.234 and raises AUC
from 0.932 to 0.951. The improvement confirms that lexical conditioning is % Full form intentional; see DECISIONS.md 2026-07-08.
not a minor residue after length, animacy, definiteness, pronominality, and
accessibility are included.

Partial pooling reaches almost the same held-out performance with far fewer
fixed-effect terms: the hierarchical model's test log loss is 0.233 and its AUC
is 0.952. It estimates 18 fixed-effect terms plus a single verb-intercept
variance, instead of one coefficient for each of the 75 verbs. The estimated verb
random-intercept standard deviation is 2.106, so the lexical signal remains
large under partial pooling.

Negative-control and simulation checks behave as they should.
Scrambled-label fits lose predictive value. Fake-data fits with no verb
signal drive the estimated verb variance to zero, while a fake verb-effect check
recovers a nonzero one. These checks
make the production result more credible, but they don't change its
explanatory level.

The result supports Bresnan et al.'s production-choice conclusion. By itself,
the same-data analysis shows reproducibility and statistical stability. What it
can't show is whether the pattern helps once the model leaves the corpus that
produced it.
