\section{Generalization and acceptability}

The BNC2014 analysis asks whether the production profile travels across corpus
design, variety, period, and annotation regime. The transport model is trained
on the spoken \texttt{languageR} rows restricted to the six BNC verbs. It's
then evaluated on harmonized BNC2014 rows.

Harmonization is lossy. The outcome maps cleanly: \texttt{VNN} becomes NP and
\texttt{VNPP} becomes PP. Verb also maps cleanly after restricting the
\texttt{languageR} data to the six shared verbs. Length is more delicate:
BNC2014 releases character lengths, while \texttt{languageR} uses token-like
lengths, so the transport script derives BNC word-count proxies in memory.

Several predictors are available only as proxies. BNC2014 has recipient and
theme animacy, theme definiteness, and recipient/theme pronominality, but some
fields have missing values and different annotation paths. Recipient
definiteness is not released as a field, so it enters only as a labelled regex
proxy in a sensitivity check. Discourse accessibility is unavailable and is
dropped from transport.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.48\linewidth}p{0.22\linewidth}rr}
\toprule
Model & Test data & Log loss & AUC \\
\midrule
\texttt{languageR} marginal & BNC2014 & 0.473 & 0.500 \\
\texttt{languageR} core transport & BNC2014 & 0.308 & 0.906 \\
\texttt{languageR} core transport & Scrambled BNC2014 & 0.885 & 0.515 \\
BNC2014 marginal holdout & BNC2014 & 0.481 & 0.500 \\
BNC2014 native core holdout & BNC2014 & 0.202 & 0.960 \\
BNC2014 native scrambled train & BNC2014 & 0.495 & 0.440 \\
\bottomrule
\end{tabular}
\caption{First BNC2014 transport and native holdout comparisons.}
\label{tab:bnc-transport}
\end{table}

The transported core model beats the marginal baseline by a wide margin. Its
test log loss is 0.308, compared with 0.473 for the \texttt{languageR} marginal
baseline, and its AUC is 0.906. When the BNC outcome is scrambled, the same
predictions collapse to 0.885 log loss and 0.515 AUC.

The BNC-native model is still stronger. On a verb-stratified BNC holdout split,
the native core model reaches 0.202 log loss and 0.960 AUC. This gap matters.
The transported model has learned a profile that's genuinely useful in the
new corpus, but the BNC data also contains local structure that the
\texttt{languageR} fit does not capture.

Several directions travel. Recipient length is negative for NP realization;
theme length is positive; recipient pronominality is strongly positive; theme
pronominality is strongly negative; \texttt{sell} and \texttt{send} are less
NP-like than \texttt{give}. These are the strongest signs that the Bresnan
profile is projectible beyond one packaged dataset.

Some details do not travel cleanly. The transported model underpredicts NP
realization for \texttt{send}, \texttt{offer}, and \texttt{lend}, and
overpredicts it for \texttt{sell}. Theme definiteness is the clearest
coefficient mismatch: negative in the \texttt{languageR}-trained model, positive
and weak in the BNC-native model.

The recipient-definiteness proxy should stay a sensitivity check. Adding it
worsens transport log loss from 0.308 to 0.347 while leaving AUC essentially
unchanged. That pattern fits the harmonization note: the proxy is interpretable,
but it should not carry the headline result.

Acceptability evidence would answer a different question. DAIS and related
judgement datasets can test whether human preferences over constructed pairs
align with the production profile. They should enter only if the paper needs a
bridge from production probability to acceptability preference. They should not
be folded into the corpus model as though they measured the same outcome.
