\section{Grammatical possibility}

The empirical result supports a sharper vocabulary, not a shortcut from
frequency to grammar. The \texttt{languageR} analysis shows that a classic
production profile is reproducible under explicit validation. The BNC2014
analysis shows that much of that profile travels to a later spoken British
corpus. Both results concern production choice.

Production choice is not the same as acceptability preference. A corpus token
records what a speaker or writer did in a context. A judgement task asks how
speakers compare alternatives under controlled conditions. These measures can
align, but alignment is an empirical result rather than a definition.

The opportunity set matters. This corpus evidence is conditioned on
attested dative-alternation tokens: cases already annotated as NP or PP
recipient realizations. It can support claims about which realization speakers
chose inside that frame. It cannot by itself estimate how often speakers might
have used a non-dative paraphrase, left the event implicit, or accepted the
less frequent member of a constructed pair.

Grammatical possibility is a third target. It concerns whether a construction
belongs to the grammar, including cases that are rare, strongly conditioned, or
unattested in a finite corpus. Production data can support claims about that
target only through an explicit bridge from attested choices to possible
alternatives.

The dative alternation makes the bridge visible. A low-probability double-object
realization may be grammatical but dispreferred under the observed predictors.
A prepositional realization may be likely because the theme is short and the
recipient is long. A form may be absent because the corpus lacks the relevant
verb, register, or discourse setting.

The current analyses therefore license a cautious claim. Bresnan et al.'s
production-conditioning profile is durable and partly transportable. That
supports the idea that corpus probabilities can reveal structure that informal
judgements miss. It does not make production probability identical to
grammatical possibility.
