\section{Conclusion}

The reanalysis has two production results and one acceptability bridge. First,
the \texttt{languageR} profile survives baselines, negative controls,
simulation checks, and partial pooling. The hierarchical model matches the
fixed-verb model with a smaller fixed-effect parameterization, so modern
modelling confirms the classic production-choice conclusion.

Second, the profile partly transports. A spoken six-verb model trained on
\texttt{languageR} predicts BNC2014 complete-case rows well above a marginal
baseline when scored on the same rows as a target-native out-of-fold benchmark.
The strongest portable signal is structural rather than verb-only: the
source-trained non-verb model nearly matches the source full model by log loss.
The BNC2014-native model still performs better. The diagnostic verdict is
bounded. Transport survives the null comparison in the high-frequency shared
core, but the target corpus is internally structured and the comparison remains
too bundled to identify separate variety, period, register, annotation, or
missingness effects.

The bridge is DAIS. On the six verbs shared with the production core, human
double-object preference tracks the production profile moderately, but the two
channels diverge in diagnostically useful places. The item and
participant-level checks reinforce that reading: after recipient condition,
theme type, and verb are included, the production score has little incremental
association with individual judgements. The \mention{offer} cases show why
acceptability evidence cannot be replaced by a larger production corpus.

That combination is the paper's centre. The classic result is not overturned.
It's made reproducible, tested against real nulls, and moved into a second
corpus where it can succeed or fail. The success is strong enough to matter,
and the failures are specific enough to guide the next analysis.

The conceptual lesson is the same one that motivates the title. Production
probability is evidence about grammatical possibility, not a synonym for it.
Open data, transport tests, and the DAIS bridge make that evidence stronger,
but they also mark a structural boundary: a production model estimates choice
within the attested frame. Additional positive-token data under the same
denominator cannot identify the grammatical-but-unattested region by itself.
Crossing that boundary needs a broader opportunity set, an explicit
extrapolative model, or another evidence channel.
